### 2024-07-11 conteudo/background/codeduplication.tex
add initial ethnography
Small revision of the text

The state-of-art methods for code clone detection usually have high computational time and memory space usage, as well being
programming language specific approches, which cannot be easily implemented in programming languagues that were not used in 
their research \citep{litreview}. These are main concern for our research and for this reason we will use a generalist textual-based 
approach based solely on Natural Languague Processing (NLP) text similarity detection methods, which we will introduce and compare
with the literature approches later on this document.

@@

\label{subsec:codumethods}

>>>>

The state-of-art methods for code clone detection usually have high computational time and memory space usage, as well being
programming language specific approches, which cannot be easily implemented in programming languagues that were not used in 
their research \citep{litreview}. These are concerns for our research, for this reason we will use a generalist textual-based 
approach based solely on Natural Languague Processing (NLP) text similarity detection methods, which we will introduce and compare
with the literature approches later on this document.

@@

\label{subsec:codemethods}

### 2024-07-11 conteudo/background/nlpsimilarity.tex
Small revision of the text

The standard methodology for text similarity detection are based on building vector embeddings of
the texts and apply a distance function in these embeddings to measure the similarity between them. 
There are multiple methods for building vector embeddings, such as bag-of-words, TF-IDF, LSI, Word2Vec \citep{gensimlivro}. 
The recent works related to build vector embeddings uses Largue Language Models (LLM), for example the usage of BERT, ELMo 
and GPT \citep{llmsimilar}. The most popular distance function used by the literature are the cosine similarity 
defined by the formula \citep{cosineref}:

@@

It is a famous fact that LLM have high computational cost, one example of such is a experiment done by Reimers and Gurevych for 
finding the most similar pair in a collection of 10,000 sentences, that required about 65 hours with BERT \citep{bertsimilar}. This 
is a similar magnitude of comparation pairs we want to do through our work and for this reason we chose to not use this kind of 
vector embedding method, but it is a valid solution for smaller codebases.

>>>>

The standard methodology for text similarity detection are based on building vector embeddings of
the texts and apply a distance function in these embeddings to measure the similarity between them. 
There are multiple methods for building vector embeddings, such as bag-of-words, TF-IDF, LSI, Word2Vec \citep{gensimlivro}. 
The recent works related to build vector embeddings uses Large Language Models (LLM), for example the usage of BERT, ELMo 
and GPT \citep{llmsimilar}. The most popular distance function used by the literature are the cosine similarity 
defined by the formula \citep{cosineref}:

@@

It is a famous fact that LLM have high computational cost, one example of such is a analyse done by Reimers and Gurevych for 
finding the most similar pair in a collection of 10,000 sentences, that required about 65 hours with BERT \citep{bertsimilar}. This 
is a similar magnitude of comparation pairs we want to do through our work, and for this reason we chose to not use this kind of 
vector embedding method, but it is a valid solution for smaller codebases.

### 2024-07-11 conteudo/tool/architecture.tex
Small revision of the text

The Setup step is composed of two components, the Function Breaker and the Code Duplication Finder. the pipeline of this step work as
follow: the Function Parser receives the codebase we are interested, extract the functions of the codebase along with metadata and
creates a new temporary codebase where it function becomes a new code file. The Code Duplication Finder iterate over every pair of 
files in the temporary codebase computing the code duplications and save them in the Code Duplication Database, which is a text file
that stores every code duplication as a triple <function1,function2,similarity>, where function1 and function2 are the functions that 
are a duplication of each other, and similarity is the metric given by the code duplication method utilized, which in our case, it is
the cossine similiratiy explained in \ref{sec:similarity}. Finally, the Query Responder consumes the temporary codebase and the 
Code Duplication Database to extract code duplication related information as per user request. A diagram to ilustrate the pipeline
can be found in the Figure (ADD FIGURE HERE). There is a explanation of how each component work below.

@@

The input codebase is expected to be a folder in the machine the tool is running into. All files in the codebase that is not a source
code file from a programming language supported by the tool will be ignored.

@@

About the programming languague's function extractor, we approached this problem in a way that any person comfortable with a language
can adapt our approach to their specific programming language. Usually, the code blocks can be easily extracted from 
source code files, independently if the code blocks are defined by curly brackets or by indentation, and it is possible to
infer if it represents a function's body given how deeply it is in the code context and a few lines before the start 
of the code block. We followed exactly this approach to implement the function extractor for the languagues we implemented. 
As a alternative approach, we can build the syntax tree \citep{compiler} of the programming language and use it to 
extract the functions, which is a more complex task then our main approach. For the programming languages we support, we implemented
it as follow:

\begin{itemize}
	\begin{item}
		\textbf{C}:
	\end{item}
	\begin{item}
		\textbf{Java}:
	\end{item}
\end{itemize}

@@

For the code duplication detection method, we visualize the source code files as texts and execute the TF-IDF vector embedding method
implemented by the Gensim library \citep{gensim} and compute the cosine similarity as the similarity metric. We chose this method 
given the alegated Gensim statements about itself performance, being a programming language independent method and the nonnecessity of 
compilable code, which is expected to happen in the temporary codebase as it not stores complete code artifacts. Platis implemented
this code cuplication detection method and distributed it to the public with a MIT license \citep{platistool} \citep{mitlicense}, 
we utilize his implementation as it has some nice to have components such as removing code comments, we do the appropriated changes
in the input and output of his implementation to respect the expected format of our tool. 
It is feasible to change the duplication detection method in a manner similar to what we did with Platis's implementation.

@@

The Setup step is a procedure executed by the user once per codebase. The Setup step receives the input codebase and the minimum
similarity metric value to a pair of functions to be considered a duplication, then it executes the Function Parser and the Code 
Duplication Finder, to generate the temporary codebase and the Code Duplication Database, which will be later used by the Query
Responder step. The value of the minimum similarity metric is to reduce the Code Duplication Database, which optimize the
memory usage and the computational cost of the Query Responder step. For example, if this parameter does not exist and we work in 
a codebase of $10000$ functions and supposing each relative path plus the function name have $50$ charactes, the Code Duplication 
Database will be a file of size approximately $5000^2 \times 50 ~= 5$ Gigabytes, as it is not expected to a real large codebase 
to the majority of pairs of functions be a duplication of each other, this number can be reduced dramastically.

>>>>

The Setup step is composed of two components, the Function Breaker and the Code Duplication Finder. The pipeline of this step work as
follow: the Function Parser receives the codebase the user are interested, extract the functions of the codebase along with metadata 
and
creates a new temporary codebase where the functions extracted becomes news code files. 
The Code Duplication Finder iterate over every pair of 
files in the temporary codebase checking if they are a code clone and if so, save them in the Code Duplication Database, 
which is a text file
that stores every code duplication as a triple \textbf{<function1,function2,similarity>}
, where function1 and function2 are the functions that 
are a duplication of each other, and similarity is the metric given by the code duplication detection method utilized, which in our case, 
it is
the cossine similarity explained in \ref{sec:similarity}. Finally, the Query Responder consumes the temporary codebase and the 
Code Duplication Database to extract duplicated functions related information as per user request. A diagram to ilustrate the pipeline
can be found in the Figure \ref{fig:diagram}. There is a explanation of how each component work through this section.

\begin{figure}
\includegraphics{diagrama_mestrado}
\caption{The architecture diagram demonstrating the relationship between the components.}
\label{fig:diagram}
\end{figure}

@@

The input codebase is expected to be a folder in the machine the tool is running into. All files in the codebase that is not a source
code file from a programming language supported by the tool will be ignored. The tool validates that a source file is supported through
analyses of the file extension.

@@

About the programming languague's function extractor, we approached this problem in a way that any person comfortable with a language
can adapt our approach to their specific programming language. Usually, the code blocks can be easily extracted from 
source code files, independently if the code blocks are defined by curly brackets or by indentation, and it is possible to
infer if it represents a function's body given how deeply it is in the code context and a few lines before the start 
of the code block. We followed exactly this approach to implement the function extractor for the languagues we implemented. 
As a alternative approach, we can build the syntax tree \citep{compiler} of the programming language and use it to 
extract the functions, which is a more complex task then our main approach. For the programming languages that e support, 
we implemented them as follow:

\begin{itemize}
	\begin{item}
		\textbf{C}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in
		the code context and some of code lines before the definition of the code block. By the C programming language's grammar, 
		the functions are defined in a global context, with depth of 0. There are other elements with the same depth as functions,
		we differentiate them by analysing if the first non-empty character before the curly brackets of the code block is a close 
		parenthesis, since functions are the only elements that respect this pattern in the C programming language.
	\end{item}
	\begin{item}
		\textbf{Java}: Given a source code file, we extract every code block from it, maintaining the depth of the code block in
		the code context and some of code lines before the definition of the code block. By the Java's grammar, 
		the functions are defined inside classes. There are different depths which we can have functions, as Java accepts inner
		classes. We differentiate functions of other elements by analysing if the first non-empty character before the curly brackets of the code block is a close 
		parenthesis. There is a problem with this approach as respect the pattern we defined for functions such as for loop, 
		and for this reason, it is required to apply other filters.
		We chose to only analyse functions of depth 1, which ignores every others elements that respect are code blocks and have
		a parentesis as the first non-empy character before the curly brackets. This choice has a counterpart, 
		functions declared inside Inner Classes are not supported by our extractor. We made this choice
		as there are not inner classes on the BigCloneBench \citep{bigclonebench}, which is the only codebase in Java explored in this work.

@@

For the code duplication detection method, we visualize the source code files as texts and execute the TF-IDF vector embedding method
implemented by the Gensim library \citep{gensim} and compute the cosine similarity as the similarity metric. We chose this method 
given the alegated Gensim statements about itself performance, being a programming language independent method and the nonnecessity of 
compilable code, which is expected to happen in the temporary codebase as it not stores complete code artifacts. Platis implemented
this code cuplication detection method and distributed it to the public with a MIT license \citep{platistool,mitlicense} ,
we utilize his implementation as it has some nice to have components such as removing code comments, we do the appropriated changes
in the input and output of his implementation to respect the expected format of our tool. 
It is feasible to change the duplication detection method in a manner similar to what we did with Platis's implementation.

@@

\label{subsec:setup}

The Setup step is a procedure executed by the user once per codebase. The Setup step receives the input codebase and the minimum
similarity metric value to a pair of functions to be considered a duplication, then it executes the Function Parser and the Code 
Duplication Finder, to generate the temporary codebase and the Code Duplication Database, which will be later used by the Query
Responder step. The value of the minimum similarity metric is to reduce the Code Duplication Database, which optimize the
memory usage and the computational cost of the Query Responder step. For example, if this parameter does not exist and we work in 
a codebase of $10000$ functions and supposing each relative path plus the function name have $50$ charactes, the Code Duplication 
Database will be a file of size approximately $5000^2 \times 50 ~= 5$ Gigabytes, as it is not expected in a real large codebase 
to the majority of pairs of functions be a duplication of each other, this number can be reduced dramastically.

### 2024-07-11 conteudo/tool/functionalities.tex
Add base figures know in a acceptable manner, 20/06
add initial ethnography
Small revision of the text

\section{Tool functionalites}
\label{subsec:func}

We proposes three main functionalities for the user in the tool, which are named duplication explorer, function information
and duplication relatory. Each of the functionalities recevies their adequated parameters to execute, with a exception
of a optional parameter shared by all of them, which is minimum similarity per query. The minimum similarity per query is different
from the siminum similarity presented in \ref{subsec:setup}, as it enables a volatile change in the minimum similarity
interested in a query specific context. There is a explanation of each of the main functionalities throughout this section.

@@

The duplication explorer is the principal functionality of our tool, the functionality purpose is to expose for the user the code duplication
pairs found by our tool. We implement optional filters methods to enable more complex queries, which are described below. A example
of the functionality usage can be found in the figure \ref{fig:explorer_ex}.

@@

	\begin{item}
		\textbf{Pattern filter:} This filter make the explorer ignore every pair of duplicated functions that does not match a pattern
		given by the user as a parameter. We say that a function matches a pattern if the string formed by the concatenation of
		the relative path of the code file plus the function name. The user can define if both functions or at least one of the 
		functions in the pair need to match the pattern by given a specific parameter.
	\end{item}
\end{itemize}

\begin{figure}
\includegraphics{explorer_example}
\caption{Explorer Example}
\label{fig:explorer_ex}
\end{figure}

@@

The Function information purpose it to give a more in depth information about a specific function. The functionality inform the user
about the relative path, the function name and the lines the function exist in the code file that the function is contained for the
function that the user is interested and every function that is a duplication of it. As there ca nbe multiple functions with the 
same name in a codebase, we distinguish then by the user providing a pattern string and we use the first function find that match
the pattern. We say that a function matches a pattern if the string formed by the concatenation of the relative path of the 
code file plus the function name. 
A example of the functionality usage can be found in the figure \ref{fig:function_ex}.

\begin{figure}
\includegraphics{function_example}
\caption{Function Example}
\label{fig:function_ex}
\end{figure}

@@

The duplication relatory purpose is to give a overview of the input codebase. The functionality extract the information
of how much lines of duplicated code is found per folder in the codebase and present it to the user in a reasonable format. To extract
the information, we use the Code Duplication Database to list the code duplications, and the temporary code to know how much lines of
duplicated code contains in each of the functions. (DO I TALK ABOUT THE TRIE USAGE TO MAINTAIN THE HIERARCHICAL TREE STRUCTURE? I 
THINK NOT). A example of the functionality usage can be found in the figure \ref{fig:relatory_ex}.

\begin{figure}
\includegraphics{relatory_example}
\caption{Function Example}
\label{fig:relatory_ex}
\end{figure}

>>>>

\section{Tool functionalities}
\label{subsec:func}

We proposes three main functionalities for the user in the tool, which are named duplication explorer, function information
and duplication relatory. Each of the functionalities receives their adequated parameters to execute, with a exception
of a optional parameter shared by all of them, which is minimum similarity per query. The minimum similarity per query is different
from the siminum similarity presented in \ref{subsec:setup}, as it enables a volatile change in the minimum similarity
interested in a query specific context. There is a explanation of each of the main functionalities through this section.

@@

The duplication explorer is the principal functionality of our tool, the functionality purpose is to expose for the user the 
duplicated functions pairs found by our tool. We implement optional filters methods to enable more complex queries, 
which are described below. A example of the functionality usage can be found in the figure \ref{fig:explorer_ex}.

@@

	\begin{item}
		\textbf{Pattern filter:} This filter make the explorer ignore every pair of duplicated functions that does not match a pattern
		given by the user as a parameter. We say that a function matches a pattern if the string formed by the concatenation of
		the relative path of the code file plus the function name contains the pattern as a substring. 
		The user can define if both functions or at least one of the 
		functions in the pair need to match the pattern by given a specific parameter.
	\end{item}
\end{itemize}

\begin{figure}
\includegraphics[scale=0.5]{explorer_example}
\caption{Explorer Example}
\label{fig:explorer_ex}
\end{figure}

@@

\label{subsec:functioncommand}

The Function information purpose it to give a more in depth information about a specific function. The functionality 
receives a function the user is interested in and then inform the user
about the relative path, the function name and the lines the function is defined for the function the user is interested and for
every function that is a duplication of the given function.
There can be multiple functions with the 
same name in a codebase, to distinguish them, we receive a pattern string from the user, and  we use the first function find that match
the given pattern. We say that a function matches a pattern if the string formed by the concatenation of the relative path of the 
code file plus the function name contains the pattern as a substring. 
A example of the functionality usage can be found in the figure \ref{fig:function_ex}.

\begin{figure}
\includegraphics[scale=0.5]{function_example}
\caption{Function Example}
\label{fig:function_ex}
\end{figure}

@@

The duplication relatory purpose is to give a overview of the input codebase. The functionality extract the information
of how much lines of duplicated code is found per folder in the codebase and present it to the user in a reasonable format. To extract
the information, we use the Code Duplication Database to list the duplicated functions, and the temporary code to know how much lines of
duplicated code contains in each of the functions. (DO I TALK ABOUT THE TRIE USAGE TO MAINTAIN THE HIERARCHICAL TREE STRUCTURE? I 
THINK NOT). A example of the functionality usage can be found in the figure \ref{fig:relatory_ex}.

\begin{figure}
\includegraphics[scale=0.5]{relatory_example}
\caption{Function Example}
\label{fig:relatory_ex}
\end{figure}

### 2024-07-11 conteudo/tool/tool.tex
Small revision of the text

Our proposed tool is a Command Line Interface (CLI) software, the tool purpose is to enable developers to query 
information about code clone duplication in a function level, by function level we mean to find functions in a codebase that 
are code duplications. The tool is separated in two steps. The 
first one, denoted as Setup, is executed only once for each codebase and it is responsible for processing the
codebase doing tasks such as extracting functions of the code files 
and searching for code clone duplications. The second step, denoted as Query Responder, is
responsible to execute the tool functionalities using the information obtained in the Setup step. More information about the 
Setup and Query Responder behavior can be found in \ref{subsec:architecture} and in \ref{subsec:func}.

>>>>

Our proposed tool is a Command Line Interface (CLI) software, the tool purpose is to enable developers to query 
information about code clone duplication in a function level, which means to find pairs of functions in a codebase
that where the functions in the pair are code clones of each other. The tool is separated in two steps. The 
first one, denoted as Setup, is executed only once for each codebase and it is responsible for processing the
codebase doing tasks such as extracting functions of the code files 
and searching for code clone duplications. The second step, denoted as Query Responder, is
responsible to execute the tool functionalities using the information obtained in the Setup step. More information about the 
Setup and Query Responder behavior can be found in \ref{subsec:architecture} and in \ref{subsec:func}.
